Generating Images From Captions\\ With Attention

### 2015-10-21 iclr-paper-new.tex
new changes

\section{Experiments}
\subsection{MNIST With Captions}
As a toy experiment, we have trained our model on the MNIST dataset with captions constructed on the fly. The $60 \times 60$ image consisted of either two or one digits, where one digit was placed in one of the four (top-left, top-right, bottom-left or bottom-right) corners of the image and two digits were either placed horizontally or vertically in nonoverlapping fashion. While most of the generative models were trained on the binarized version of MNIST dataset, our model was trained directly on pixel intensities with binary cross entropy cost function. The generated images and the attention alignment are displayed in Figure

@@

Aside from flipping words in the captions that describe the common situations, we were curious about trying nonsense captions that model has never seen in the training phase. These captions described some object, like ``elephants'', doing something in the completely new environment, like ``flying in the blue skies''. That test shows whether the model truly generalizes to completely new situations or overfits on the scenes that the model seen during training. The model managed to generalize to new situations, as shown in Figure~\ref{fig:genimages4}.

@@

In Figure~\ref{fig:diffmodels}, we generated several samples from the prior of each of the current state-of-the-art generative models, conditioned on the caption ``A group of people walk on a beach with surf boards". While all of the samples look sharp, the images generated by LAPGAN look more noisy and it is harder to make out definite objects, 
%don't have a clear structure in them
whereas the images generated by variational models trained with L2 cost function have a watercolor effect on the images.

As for the quantitative comparison of different models, we first compare the performances of the model trained with variational methods. We rank the images in test set conditioned on the captions based on the variational lower bound of likelihood and then report the Precision-Recall metric to evaluate the quality of the generative model. Unsurprisingly, the quality of image retrieval using generative models is worse than that of discriminative models that were specifically built for retrieval. To deal with the large computational complexity involved in looping through each test image, we create a shortlist of one hundred images including the correct one, based on the images having the closest distance in the convolutional feature-space of a VGG-like model trained on the CIFAR dataset. Since there are ``easy'' images for which the model assigns high likelihood independent of the query caption, we look at the ratio of likelihood of image conditioned on the sentence to likelihood of image conditioned on the mean sentence representation in the training set. We found that the reconstruction error $\rloss$ increased for sharpened images, and that sharpening considerably hurt the retrieval results. Since sharpening changes the statistics of images, computing reconstruction error for each pixel is not necessarily a good metric.

Instead of calculating error per pixel, we turn to a smarter metric, the Structural Similarity Index (SSI), which incorporates luminance and contrast masking into the error calculation. Due to strong inter-dependencies of closer pixels, the metric is calculated on small windows of the image. To calculate SSI, we sampled fifty images from the prior of each generative model for every caption in the test set.

@@

>>>>

\section{Experiments}
\subsection{MNIST With Captions}
As a first experiment, we trained our proposed model on the MNIST dataset with captions constructed on the fly. Either one or two digits from MNIST training dataset were placed on the $60 \times 60$ blank image. One digit was placed in one of the four (top-left, top-right, bottom-left or bottom-right) corners of the image. Two digits were either placed horizontally or vertically in nonoverlapping fashion. While most of the generative models were trained on the binarized version of MNIST dataset, our model was trained directly on pixel intensities with binary cross entropy cost function. The generated images and the attention alignment are displayed in Figure~\ref{fig:figmnist}.

@@

As a real test of model generalization, we were tried generating images corresponding to captions that describe scenarios that never occur in real life. These captions describe some common object doing unusual things considering the physics of real world. Even though these scenarios may never happen in real life, it is very easy for humans to imagine the corresponding scene. Nevertheless, as you can see in Figure~\ref{fig:genimages4}, the model managed to generate reasonable images.

@@

We compared the performance of the proposed model to the DRAW model conditioned on captions without align function (noalignDRAW) as well as the DRAW model conditioned on skipthought vectors (skipthoughtDRAW). All of the conditional DRAW models were trained with binary cross entropy cost function. We also compared our model with Fully-connected (FC) and Convolutional-Deconvolutional (Conv.) Variational Autoencoders which were trained with the least squares cost function. The LAPGAN model was trained on two level Laplacian Pyramid and was conditioned on the skipthought vectors.

In Figure~\ref{fig:diffmodels}, we generated several samples from the prior of each of the current state-of-the-art generative models, conditioned on the caption ``A group of people walk on a beach with surf boards". While all of the samples look sharp, the images generated by LAPGAN look more noisy and it is harder to make out definite objects, whereas the images generated by variational models trained with L2 cost function have a watercolor effect on the images.

As for the quantitative comparison of different models, we first compare the performances of the model trained with variational methods. We rank the images in test set conditioned on the captions based on the variational lower bound of likelihood and then report the Precision-Recall metric to evaluate the quality of the generative model (see Table~\ref{tab:results}.). Unsurprisingly, the quality of image retrieval using generative models is worse than that of discriminative models that were specifically built for retrieval. To deal with the large computational complexity involved in looping through each test image, we create a shortlist of one hundred images including the correct one, based on the images having the closest distance in the convolutional feature-space of a VGG-like model trained on the CIFAR dataset. Since there are ``easy'' images for which the model assigns high likelihood independent of the query caption, we look at the ratio of likelihood of image conditioned on the sentence to likelihood of image conditioned on the mean sentence representation in the training set\cite{kiros_captions}. We found that the reconstruction error $\rloss$ increased for sharpened images, and that sharpening considerably hurt the retrieval results. Since sharpening changes the statistics of images, computing reconstruction error for each pixel is not necessarily a good metric.

Instead of calculating error per pixel, we turn to a smarter metric, the Structural Similarity Index (SSI), which incorporates luminance and contrast masking into the error calculation. Due to strong inter-dependencies of closer pixels, the metric is calculated on small windows of the image. Due to independence property of test captions, we sampled fifty images from the prior of each generative model for every caption in the test set in order to calculate SSI. The results are reported in Table~\ref{tab:results}.

@@

\section{Discussion}

We have shown that our proposed model manages to generate reasonable images corresponding to respective captions as well as generalize to completely novel scenarios that never occur in real life. Sequential generation of images using attention played a crutial part in it. However the DRAW model still underfits on training data which leads to the blurry samples. We think that extending this process by generating very small sections of images like superpixels over potentially hundreds of timesteps might help the model fit the data and generate sharp images.

\section{Acknowledgements}
We would like to thank developers of Theano for a powerful software and authors of \cite{denton_lapgan} for open sourcing their code. We would also like to thank Ryan Kiros and Nitish Srivastava for helpful discussions. We will release our code with pretrained models as soon as we clean it. Acknowledging funding (Russ)
